---
title: "A3: SPL Library Checkouts"
output: html_document
---

```{r, include=FALSE}
source("summary.R")
```

### Introduction

The data set *Checkouts-SPL-Data* from 2017 to 2023 with at least 10 times checkouts is from the Seattle Public Library. I choose this data set because I want to explore people reading preference, such as what books are selected most or published in which year, among popular books that are filtered by the at least of 10 times of checkouts. And then, I want to manipulate the data to analyze their habits of reading, such as how long spent on one book, what do they use to read. Since reading is good habit to have in life, if more people can enjoy reading, it will be a good way to spread different culture, ideas, and knowledge through books. In order to make the communication of words more effective, understanding people's preferences and habits of reading provides a foundation for technological company and publishing house to better target consumers.

### Summary Information

There are six question I want to explore in this project.

**What is the name of the book that was read the most times in a month?**

The most popular book in a month is named `r df.list$max.checkouts.name`.

**Which book was read the most times in a month?**

It has been checked out by `r df.list$max.checkouts` times within that month.

**In what year do people most like to read books published?**

People like to read books that were published in `r df.list$popular.year`.

**How long is the average borrowing/reading period for people?**

This question cannot be answer by given data set since we do not have data recording how long did each book being hold and how fast did it change hand.

**What is the difference between the number of people who like to read paper books and e-books?**

From 2017 to 2023, `r df.list$dif.ebook.book` more people prefer reading paper book than ebook.

**Which checkout type is the most popular among these years?**

The most popular check out type is `r df.list$type`.

By exploring the questions above, we can have a basic understanding about people's reading preference and thus save time once using the data set. Since we know there is no information about how long does one holding the book, we wouldn't plot the data for this question. Besides, We realize that the most popular vendor is Horizon, we may expect using a pie chart to visulaize the proportion of each vendor, and checking whether Horizon takes the largest part. People also like the book published in 2017, we can do some reseach and see why does the books published in the year is very popular, and gain more insight about that. And also, other values explored above can be use to better utilize the data and create better visulizations.

### The Dataset

The data is collected by Seattle Public Library. It includes a monthly count of checkouts by title for physical and electronic items. The dataset starting info is from in April 2005, but we use part of the data set which includes the data from 2017 to 2023. Within the data set, we have variables of "UsageClass", denoting if item is physical or digital; "CheckoutType", indicating the vendor tool; "CheckoutYear", year that books were checked out; "Title", the full name of the book, etc.

Checkout data is from different current and historical sources. For example, digital items, the media vendors: Overdrive, hoopla, Freegal, and RBDigital provide usage data. The data was collected for the purpose of increasing the quality of life for our residents by providing more transparency, accountability and comparability information, which will also give foundation for economic development,research utilization, and internal management.

We should consider whether the data include personal information that entails privacy. For example, Readers' real borrowing names and preferences should not be disclosed to the public. That's exactly what the SPL is doing, hiding the names of its readers to protect private privacy at the same time provide data to the public.

However, this data set does not include the data about how long did people holding the book. It is hard to understand the time interval that the same book change hand.

### First Trends Over Time Chart

```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart1.R")}
```

The first plot is a stacked bar chart, the taller a bar is, the larger the volume of those numeric values. The bars are stacked showing how each material type comparing to others over time. 

With the development of technology, people today embrace many different kinds of types to read books. Some prefer paper book, while others prefer e-book, and there are also some people like audio book as well as video book. Realizing today's people reading preference helps to use different reading tools more effectively and to deliver different information to different groups.

The stacked bar chart shows that audio book is the least popular tool for reading because given any year, it takes the smallest part of a bar. And we can also see that the popularity of video book is decreasing over time. That might indicating that in the future, video book will be abandoned. Besides, book and ebook take the most part of a bar even though in 2020 the amount of people choosing book fluctuated to lower level. We can see that after 2020, preference of book increases, and since 2017, preference of ebook has increased. We may expect that in the future, these two types of material will be the most popular two.

### Second Trends Over Time Chart

```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart2.R")}
```

I choose to plot time series data to visualize the trend of the amount of checkouts over time with respect to the usage class, digital or physical. By ploting a time series data, we can easily see how does people preference towards the medium for reading change. It is very important to know whether digital books are preferred than physical books, and library can utilize that information to change the book storage and appropriatly modify paper and electronic books proportion.

The graph shows that there is huge down in physical book around 2020. This significant drop of reading physical books may be contributed to the outbreak of pandemic; people did not often go out to the libraries and buying or reading physical books there but instead shopping online and read online books. Besides, we may expect that a huge drop in the choice of physical books would lead to a huge increase in digital books around 2020, though we see digital books increases at that time but it does not icreases that huge amount. There must be some reasons account for the lost difference between the two classes. However, from a macro-scale, seeing the upper and upper red line, we realize that more and more people choose digital books. Before 2020, physical books shows a decreasing trend, which is aggravated by the huge downturn in 2020. It gradually recovers at the end of 2020 but the total amount is still lower than the digital.

### Box Plot

```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart3.R")}
```

The third plot I choose to use box plots. Box plot is good at telling the mean, max, min of the data, and so we can get better understanding about the distribution of the amount of choosing each vendors throughout these years. 

By looking at the plot, we can see the mean of Horizon is 7769 meaning that throughout the 10 years, about average 7769 times people choosing to use Horizon per month. We can see the mean of Horizon is higher than all other vendors', the first quantile and third quantile are also higher than these of others, meaning that Horizon is the most popular vendor among the five. And the second popular vendor is OverDrive. Besides, we can barely see the distribution of the other three vendors. That might tells us only few people choose them. 

We should be careful when consider the use of Horizon or OverDrive because though the mean of Horizon is higher, its variance and range are also higher. We can also see that the minimum of Horizon is zero, meaning that there are months when people did not choose to use Horizon. And the data of OverDrive is more concentrated, indicating stable distribution. 
